### 2020-08-19 paper.tex
Restructure Prev Efforts

@@

Our work is not the first attempt to understand and improve on the struggles software engineers have to overcome when working on RP code, either with or without RxJS. We are going to present a selection of four utilities, whereas two of them have an academic background. We learned about them during our research either because they were mentioned in one of the interviews and war stories, or because we have worked with them in the past.

Salvaneschi et al. did focus their research heavily on RP during the mid 2010s. Besides the creation of REScala in 2014 \cite{10.1145/2577080.2577083}, an RP DSL and runtime for the Scala programming language, they have looked into the debugging habits of software engineers as well. In the paper ``Debugging for Reactive Programming'' \cite{10.1145/2884781.2884815}, Salvaneschi et al. elaborate on the necessity of RP debugging tools that understand the specific language abstractions. They provide with the \emph{Reactive Inspector} plugin for the Eclipse IDE a reactive debugger for the REScala RP runtime environment. One of its main features is the visualization of the internal dataflow graph. A new class of  breakpoints, the ``reactive breakpoints,'' builds upon the visualized graph as well: They do not depend on a specific line of code or statement rather than on a specific node in the graph. The execution of the program can be interrupted e.g., once a node retains a specific output value or an event of interest takes place in the runtime. Along with the RP specific debugging features, Reactive Inspector implements omniscient debugging concepts as well: Time travel allows to record and replay runtime behavior at a later time without re-execution.

Salvaneschi et al. claim\cite{10.1145/2577080.2577083} the architecture behind \emph{Reactive Inspector} to be flexible enough to be used with other RP runtime environments.

\emph{rxfiddle} as proposed by Banken et al. is the first reference implementation of their RP debugger architecture for the ReactiveX API specification described in their paper ``Debugging Data Flows in Reactive Programs'' \cite{10.1145/3180155.3180156}. The design defines two independent parts: The (i) \emph{Host instrumentation} augments a ReactiveX API implementation to emit events at runtime (e.g., when an Observable is created or subscribed) and collects them. The (ii) \emph{Visualizer} component interprets these events and presents them along two dimensions: The StoryFlow graph \cite{YWu2013a} shows when an Observable is created and how it interacts with other Observables. A marble diagram visualizes the values emitted over time for every Observable further.

The reference implementation supports event collection for RxJS-based code in versions 4 and 5. It is available as an online application where snippets of code can be inspected. A proof-of-concept implementation which works on a local development machine is accessible via the projects Git repository.

The architectural design of \emph{rxfiddle} is similar to the one of \emph{Reactive Inspector}: Banken et al. claim that their Visualizer can be adapted to different RP runtime environments as well.

@@

Software engineers tend to augment their code with additional log trace statements where they have to track down a problem we saw earlier. rxjs-spy\cite{rxjsspy} intends to improve on this by introducing the \mintinline{TypeScript}{tag} operator, along with sophisticated log monitoring facilities. Ideally, ``tagged'' Observables are created during development when the dataflows are composed for the first time. Like the \mintinline{TypeScript}{tap} operator, a tag as shown in Listing~\ref{lst:rxjs-spy-tag} on Line~3 does not influence runtime behavior of an application.

@@

		interval(1000).pipe(
		  map(i => i * 2),
		  tag('multiply'),     // Tag Observable with "multiply"
		  map(i => i - 1),
		  take(2)
		).subscribe();
	\end{minted}
	\caption{Usage of \emph{rxjs-spy} \mintinline{TypeScript}{tag} Operator on Line~3}
	\label{lst:rxjs-spy-tag}
\end{listing}

Once necessary, rxjs-spy can be advised to log events such as subscribe, emit, error, complete and unsubscribe of matching tags directly to the console (see Listing~\ref{lst:rxjs-spy-log}) without touching any Observable code at all. The additional events logged by rxjs-spy were noted invaluable in the interviews presented earlier as well as in our own experience.

\begin{listing}[H]
	\begin{minted}{Typescript}
		Tag = multiply; notification = subscribe
		Tag = multiply; notification = next; value = 0
		Tag = multiply; notification = next; value = 2
		Tag = multiply; notification = unsubscribe
	\end{minted}
	\caption{Trace log generated by \emph{rxjs-spy} \mintinline{TypeScript}{tag} from Listing~\ref{lst:rxjs-spy-tag}}
	\label{lst:rxjs-spy-log}
\end{listing}

Additional features are available through the debuggers console interface. Engineers can ``pause'' a tagged Observable, so it collects incoming values. They can then be emitted one after another manually or resumed at once.

\subsection{Rx Visualizer}

The \emph{Rx Visualizer}\cite{rxviz} is a utility available online in the browser. It is an ``animated playground for Rx Observables''\cite{rxviz} and allows to visualize snippets of RxJS code using marble diagrams. Diagrams are generated over a given, fixed time and available for download as SVG\footnote{\url{https://developer.mozilla.org/en-US/docs/Web/SVG}} files.

The visualizer helps to understand the behavior of a specific part of an Observable stream in a sandboxed environment as a reported war story stated.

@@

@@

A graphical representation of the dataflow graph is better suited to show dependencies within the model then textual trace logs. Previous research showed that the sheer amount of nodes in an elaborate graph is hard to visualize effectively \cite{10.1145/3180155.3180156} on one side, on the other also not easily implement efficiently as well.

>>>>

Before we elaborate on a solution in order to overcome the challenges identified with RQ1, we are going to present a selection of RP specific debugging utilities available today.

@@

@@

\emph{rxjs-spy}\cite{rxjsspy} is a logging library specialized on RxJS observables. Once an observable is tagged with an arbitrary string identifier, the library can generate trace logs when a value is emitted as well as when individual lifecycle events occur (i.e. subscribe, unsubscribe, complete and error). Ideally, tagged Observables are created during development when the dataflows are composed for the first time. Like the \mintinline{TypeScript}{tap} operator in Listing~\ref{lst:rxjs-debugging}, the \mintinline{TypeScript}{tap} operator in Listing~\ref{lst:rxjs-spy-tag} on Line~3 is completely transparent to the actual dataflow.

@@

		interval(1000).pipe(
		  map(i => i * 2),
		  tag('multiply'),     // Tag Observable with "multiply"
		  map(i => i - 1),
		  tag('subtract')      // Tag Observable with "subtract"
		  take(2)
		).subscribe();
	\end{minted}
	\caption{Application of rxjs-spy using its \mintinline{TypeScript}{tag} operator on Line~9 and 11}
	\label{lst:rxjs-spy-tag}
\end{listing}

The dataflow configuration in Listing~\ref{lst:rxjs-spy-tag} will produce a trace log as shown in Listing~\ref{lst:rxjs-spy-log}.

\begin{listing}[H]
	\begin{minted}{Typescript}
		Tag = multiply; notification = subscribe
		Tag = multiply; notification = next; value = 0
		Tag = multiply; notification = next; value = 2
		Tag = multiply; notification = unsubscribe
	\end{minted}
	\caption{rxjs-spy execution trace log generated from Listing~\ref{lst:rxjs-spy-tag}}
	\label{lst:rxjs-spy-log}
\end{listing}

Additional features are available through the debuggers console interface. E.g., a tagged Observable can be paused so values get collected rather than being emitted immediately. The engineer can then emitted these values one after another manually or resume all of them at once.

\emph{rxfiddle} as proposed by Banken et al.\cite{10.1145/3180155.3180156} is the first reference implementation of their RP debugger architecture for the ReactiveX API specification. They describe a software design consisting of two independent components: The (i) \emph{host instrumentation} augments a ReactiveX API implementation to emit events at runtime (e.g., emitting a value or lifecycle events) and forwards them to the second component. The (ii) \emph{visualizer} interprets the events and displays them along two dimensions: The StoryFlow graph\cite{YWu2013a} shows when an Observable is created and how it interacts with other Observables, whereas a marble diagram visualizes the values emitted over time for every Observable.

The reference implementation supports event processing for the (outdated) RxJS major versions 4 and 5 only and is available as an online application\footnote{\url{https://rxfiddle.net/}}. A proof-of-concept implementation which can be executed on a local computer is available through the projects Git repository\footnote{\url{https://github.com/hermanbanken/RxFiddle}}.

\subsection{RxViz}

\emph{RxViz} is a visualizer utility available online\footnote{\url{https://rxviz.com/}}. It is an ``animated playground for Rx Observables''\cite{rxviz} and allows the visualization of RxJS observables using marble diagrams. Engineers implement or copy-paste dataflows in an editor window using JavaScript. This code is then used to generate a diagram over a configurable time interval. The diagrams are rendered immediately and are available as downloadable SVG files.

As of writing this paper, the projects owner stopped working actively on the open source application and is looking for a new maintainer\cite{rxviz-maintainer}.

In the paper ``Debugging for Reactive Programming''\cite{10.1145/2884781.2884815}, Salvaneschi et al. elaborate on the necessity of RP debugging tools which can interpret its specific dataflow abstractions. Based on their research, developed a reactive debugger named \emph{Reactive Inspector}. This plugin for the Eclipse IDE provides reactive debugging functionality for programs using the REScala RP runtime.

The plugins core feature is the visualization of the dataflow graph. This representation of the dataflow provides a UI to add a new kind of breakpoint, the ``reactive breakpoint,'' to individual nodes in the graph. With them, the execution of the program can be interrupted e.g., when a node retains a specific value or a particular lifecycle event takes place.

Along with the RP specific debugging features, Reactive Inspector experiments with the omniscient debugging concept: The time travel functionality allows to record dataflows. These recordings can then be replayed and inspected at a later time without actual re-execution of the program.

@@

\textbf{TODO REWRITE HERE}

@@

A graphical representation of the dataflow graph is better suited to show dependencies within the model then textual trace logs. Previous research showed that the sheer amount of nodes in an elaborate graph is hard to visualize effectively\cite{10.1145/3180155.3180156} on one side, on the other also not easily implement efficiently as well.
